{\LARGE Basis Choice for Scalar-on-Function Regression \\ with Applications to Near-Infrared Spectroscopy}

### 2022-02-08 Paper/main.tex
Karhunen Loeve Wording

	\subsection{Karhunen-Lo\'{e}ve Expansion and Empirical Eigenbases}\hypertarget{KL}{}
	Given a realization of a random function $X: \Omega \mapsto \mathbb{L}^2[0,1]$, it is possible to represent this realization in terms of its generating stochastic process. To do so, it is necessary to define the mean and covariance functions of $X(\omega)$.
	\begin{equation}\label{MeanFunction}
		\mu(t) = \mathbb{E}\left[ X(\omega)(t) \right]
	\end{equation}
	\begin{equation}\label{CovarianceFunction}
		c(t,s) = \mathbb{E}\big[ \left( X(\omega)(t) - \mu(t) \right) \left( X(\omega)(s) - \mu(s) \right) \big]
	\end{equation}

	where the $c(t,s)$ are Hilbert-Schmidt Kernels defined through $c : [0,1] \times [0,1] \rightarrow \mathbb{R}$. Let $K$ be a Hilberst-Schmidt operator on $\mathbb{L}^{2}[0,1]$ such that $K : \nu \rightarrow K \nu$ for $\nu \in \mathbb{L}^{2}[0,1]$, by
	\begin{equation}\label{HSKernal}
		[K \nu](t) = \int_{0}^{1}c(t,s) \nu(s)ds = \lambda \nu(t)
	\end{equation}
	More exact definitions are provided at \ref{def_HS}. Then, the operator $K$ has orthonormal basis functions $\nu^{m} \in \mathbb{L}^{2}[0,1]$ corresponding to eigenvalues $\lambda^{m}$ for all $m$ since it is a positive compact self-adjoint operator (cf. \cite{alexanderian_KLexpansion_2015}). Moreover, $K$ holds that the eigenvalues can be ordered in nonincreasing order as follows $\lambda^{1} \geq \lambda^{2} \geq \dots \geq 0$ where the superscript is not the power but index. Therefore, the functions $X$ are approximated enough well by first few principal components since the order of them are sorted in descending order of eigenvalues corresponding to the eigenfunctions (e.g. $Var(\xi^{m}) \geq Var(\xi^{n})$ for all $m < n$). Theoretical considerations lead to the result that $X$ can be represented in the following form, called its Karhunen-Lo\'{e}ve expansion. The proofs are provided at \ref{Proof1} and \ref{Proof2}.
	\begin{equation}\label{KarhunenLoeve}
		X(\omega)(t) = \mu(t) + \sum_{m \in \mathbb{N}} \xi^m(\omega) \nu^m(t), \quad \xi^m(\omega) =  \int_{0}^{1} \left(X(\omega)(s) - \mu(s)\right) \nu^m(s) \mathrm{d}s
	\end{equation}

	where the $\nu^m$ are defined by the countable set of solutions $\{(\lambda^m, \nu^m) \: \vert \: m \in \mathbb{N}\}$ of ({\ref{HSKernal}}). The random variables $\xi^{m}(\omega)$ satisfy following properties.
	\begin{multicols}{2}
		\begin{enumerate}
			\item $\mathbb{E}\left[\xi^m(\omega)\right] = 0$
			\item $Cov\left(\xi^m(\omega), \xi^n(\omega)\right) = 0$ if $m \neq n$
			\item $Var\left(\xi^m(\omega)\right) = \lambda^m$
		\end{enumerate}
	\end{multicols}

@@

	With these it is possible to derive a set of sample analogs $\{(\hat{\lambda}^m, \hat{\nu}^m) \: \vert \: m = 1, \dots, N-1\}$ for $\{(\lambda^m, \nu^m) \: \vert \: m \in \mathbb{N}\}$ as the solutions of the following equation. As in the case of ordinary principal components, the number of eigenfunctions corresponding to non-zero eigenvalues is limited (cf. chapter 8.2.3 \cite{ramsay_functional_2005}). As each curve is infinite dimensional, there is no upper limit to this number due to the dimensionality. However the number of curves still imposes an upper limit of $N-1$ non-zero eigenvalues, where $N$ is the number of curves in the data set.
	\begin{equation}
		\int_{0}^{1}\hat{c}(t,s)\hat{\nu}(s) \mathrm{d}s = \hat{\lambda} \hat{\nu}(t)
	\end{equation}

	This naturally leads to the following representation.
	\begin{equation}
		x_i(t) = \hat{\mu}(t) + \sum_{j = 1}^{N-1} \hat{\xi}_{i}^{m} \hat{\nu}^{m}(t)
	\end{equation}

>>>>

	\subsection{Karhunen-Lo\'{e}ve Expansion and Empirical Eigenbases}\hypertarget{KL}{}
	Given a random function $X: \Omega \mapsto \mathbb{L}^2[0,1]$, it is possible to represent its realizations in terms of the stochastic process. To do so, we define the mean and covariance functions of $X(\omega)$.
	\begin{equation}\label{MeanFunction}
		\mu(t) = \mathbb{E}\left[ X(\omega)(t) \right]
	\end{equation}
	\begin{equation}\label{CovarianceFunction}
		c(t,s) = \mathbb{E}\big[ \left( X(\omega)(t) - \mu(t) \right) \left( X(\omega)(s) - \mu(s) \right) \big]
	\end{equation}

	where the $c: [0,1] \times [0,1] \rightarrow \mathbb{R}$ are Hilbert-Schmidt Kernels and $K$ is a corresponding Hilbert-Schmidt operator\footnote{Definitions of Hilbert-Schmidt Kernels and Operators are provided in Appendix \ref{def_HS}.} $K : \nu \rightarrow K \nu$ for $\nu \in \mathbb{L}^{2}[0,1]$ defined by
	\begin{equation}\label{HSKernal}
		[K \nu](t) = \int_{0}^{1}c(t,s) \nu(s)ds = \lambda \nu(t)
	\end{equation}

	Then, the operator $K$ has orthonormal basis functions $\nu^{m} \in \mathbb{L}^{2}[0,1]$ each corresponding to an eigenvalue $\lambda^{m}$ since it is a positive compact self-adjoint operator (cf. \cite{alexanderian_KLexpansion_2015}).
	Theoretical considerations lead to the result that $X$ can be represented in the following form, called its Karhunen-Lo\'{e}ve expansion.\footnote{Proofs for these theorems are provided in Appendix \ref{Proof1} and \ref{Proof2}.} 
	\begin{equation}\label{KarhunenLoeve}
		X(\omega)(t) = \mu(t) + \sum_{m = 1}^{\infty} \xi^m(\omega) \nu^m(t), \quad \xi^m(\omega) =  \int_{0}^{1} \left(X(\omega)(s) - \mu(s)\right) \nu^m(s) \mathrm{d}s
	\end{equation}

	where the $\nu^m$ are defined by the countable set of solutions $\{(\lambda^m, \nu^m) \: \vert \: m \in \mathbb{N}\}$ of Equation \ref{HSKernal} and the random variables $\xi^{m}(\omega)$ satisfy the following properties.
	\begin{multicols}{2}
		\begin{enumerate}
			\item $\mathbb{E}\left[\xi^m(\omega)\right] = 0$
			\item $Cov\left(\xi^m(\omega), \xi^n(\omega)\right) = 0$ if $m \neq n$
			\item $Var\left(\xi^m(\omega)\right) = \lambda^m$
		\end{enumerate}
	\end{multicols}

	As in the well known setting of principal component analysis, the eigenvalues correspond to the variance in the random function that is explained by the corresponding eigenfunction. 
	Therefore, ordering the the eigenfunctions according to their their corresponding eigenvalues $\lambda^{1} \geq \lambda^{2} \geq \dots \geq 0$ is useful for approximation purposes. Due to this property a functional observation can often be approximated well by using a limited number of the eigenfunctions of its generating random process.\\

@@

	With these it is possible to derive a set of sample analogs $\{(\hat{\lambda}^m, \hat{\nu}^m) \: \vert \: m = 1,\: \dots\:, N-1\}$ for $\{(\lambda^m, \nu^m) \: \vert \: m = 1, 2, \:\dots\}$ as the solutions of the following equation. 
	\begin{equation}
		\int_{0}^{1}\hat{c}(t,s)\hat{\nu}(s) \mathrm{d}s = \hat{\lambda} \hat{\nu}(t)
	\end{equation}

	In the following, we will call the $\hat{\nu}^m(s)$ functional principal components to distinguish the sample analogs from the theoretical eigenfunctions $\nu^m(s)$. As in the case of ordinary principal components, the number of functional principal components corresponding to non-zero eigenvalues is limited (cf. chapter 8.2.3 \cite{ramsay_functional_2005}). As each curve is infinite dimensional, there is no upper limit to this number due to the dimensionality. However, the number of curves still imposes an upper limit of $N-1$ non-zero eigenvalues, where $N$ is the number of curves in the data set.
	These sample analogs naturally lead to the following representation of each sample curve $x_i(t)$.
	\begin{equation}
		x_i(t) = \hat{\mu}(t) + \sum_{j = 1}^{N-1} \hat{\xi}_{i}^{m} \hat{\nu}^{m}(t)
	\end{equation}
